---
title: "Analysis of chum salmon bycatch from shoreside sector of Bering Sea Aleutian Islands pollock trawl fishery."
author: "Patrick Barry and Wes Larson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

CrntSW <- 28 #Current Stat week
CrntYr <- format(Sys.Date(),"%Y") #Get the current year

if(!exists(file.path(paste("./Reports/",CrntYr,"/",CrntSW,sep=""))),
  dir.create(file.path(paste("./Reports/",CrntYr,"/",CrntSW,sep=""))))
```

## Summary

Main results from analysis w/ table

## Introduction

Short summary of the project, outcomes etc.
> Fishery & bycatch
> Management under NPFMC
> NOAAs current sampling & reporting
> Impetus for project
> BBSRIs project outline ~ Appendix methods

## Chum Salmon Bycatch Encounters 

Read in Glenns file
```{r}
AKRO_Files <- list.files(file.path(paste("./MainData/AKRO/",CrntYr,sep="")))

AKRO_Obs <- read.csv(paste("./MainData/AKRO/")
```

plot total bycatch by date

## Sampling of Bycatch



## Genotyping 
Read in genotypes
Bind the catch info to the genotypes
Subset to ABL markers
Perform stock comp analysis 
Expand to number of fish from catch data

Compare with prior weeks results

## Appendix I: Methods
### Genotyping
### Genetic Stock Composition Estimates
